---
title: "DTW_PCA"
author: "Katharina Hellmund & Nadia Hajighassem"
date: "2023-12-29"
output:
  html_document:
     toc: true
     toc_float: true
     number_sections: true
     theme: journal
mainfont: Times New Roman
geometry: margin=1in
knit: (function(inputFile, encoding) {
  browseURL(
    rmarkdown::render(
      inputFile,
      encoding = encoding,
      output_dir = "html.file",
      output_file = "DTW_PCA.html"))})
---

```{r setup, message = FALSE, warning = FALSE, include = FALSE}
## Setup chunk
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)


pacman::p_load(
  "XML",
  "tidyverse",
  "fs",
  "assertthat",
  "stringi",
  "dtw",
  "RTransferEntropy",
  "signal",
  "conflicted",
  "Rcpp",
  "future",
  "dtw",
  "dplyr",
  "plotly",
  "htmlwidgets",
  "hrbrthemes",
  "zoo"
)
conflicts_prefer(dplyr::filter)
## Making sure we are in the right directory
wd <- getwd()
if (basename(wd) != "PerceptionActionExam") {
  setwd("./PerceptionActionExam")
}

                              ##NADIA PATH##
  data_dir <- path_home() %>% 
   path("Documents", "GitHub", "PerceptionActionExam-Clean-up-attempt", "data", "tsvs") ## Nadia's path

                              #KATHARINA PATH##
 #data_dir <- path_home() %>%
 # path("Desktop","UNI", "3.semester", "Perception & Action", "PerceptionActionExam", "data", "tsvs") ## Katharina's path


#Aesthetic setup
theme_set(theme_ipsum(base_family = "Times New Roman"))
global_fill_colour <- "#8d5b5a"
aesthetic_palette <- c(
  "#d8aeb5","#c17f8c","#b59592","#9b6f69","#a94f62","#8d5b5a","#684141","#733545","#523438","#48222b","#2f1a1b")
aesthetic_highlight_difference_palette <- c("#d8aeb5","#2f1a1b")
```

# DTW on PCA

## Importing components

```{r}
# Directory where the components are stored
components_dir <- "results/PCA"

                  #IMPORTING ALL COMPONENTS FROM A#
# List all CSV files in the directory that include "_A" in their names
component_files_A <- fs::dir_ls(path = components_dir, glob = "*_A.csv")

component_list_A <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in component_files_A) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  component_list_A[[file_name]] <- read.csv(csv_file)
}

                  #IMPORTING ALL COMPONENTS FROM B#
# List all CSV files in the directory that include "_B" in their names
component_files_B <- fs::dir_ls(path = components_dir, glob = "*_B.csv")

component_list_B <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in component_files_B) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  component_list_B[[file_name]] <- read.csv(csv_file)
}

#IMPORTING COMPONENT WEIGHTS#
PCA_component_weights <- read_csv("results/PCAcomponentweights.csv")
```

## DTW Scoring each pair (Two components)

In creating the PCA, we also made screeplots to see how much of the data was described by each component. As so, it would not be reflective of the data if both components were weighted equally. Therefore, the weights that were defined previously have been imported, and will be used to assign a weight to each component. For the remainder of what is not explained by either component 1 or 2, the second component will inherit the weight from component 3,4,5 etc.

This means, that if component number 1 explains 70% of the data, the DTW score will be multiplied by 0.7. Component 2 will then count for the remainder (1-0.7) even if it only counts for 25%. This is done in order to spare computational power, and in most cases, two components count for \>90% of data,

First we find the DTW for each pair:

```{r}
# Load the dtw library
library(dtw)

# Initialize a dataframe to store the results
dtw_results_df_combined <- data.frame(DF_Name = character(0), # Add a column for dataframe names
                                      DTW_Component_1 = numeric(0),
                                      DTW_Component_2 = numeric(0))

# Retrieve the names of the dataframes in component_list_A
df_names <- names(component_list_A)

# Loop through each index in the component lists
for (i in seq_along(component_list_A)) {
    # Select the i-th dataframe from each list
    component_A_df <- component_list_A[[i]]
    component_B_df <- component_list_B[[i]]

    # Extract the first PCA component from each dataframe
    component_A_1 <- component_A_df[, 1] # Assuming the first column is the first PCA component
    component_B_1 <- component_B_df[, 1] # Same assumption for component_list_B

    # Calculate DTW distance for the first PCA component
    dtw_distance_1 <- dtw(component_A_1, component_B_1, keep = TRUE)$distance

    # Extract the second PCA component from each dataframe
    component_A_2 <- component_A_df[, 2] # Assuming the second column is the second PCA component
    component_B_2 <- component_B_df[, 2] # Same assumption for component_list_B

    # Calculate DTW distance for the second PCA component
    dtw_distance_2 <- dtw(component_A_2, component_B_2, keep = TRUE)$distance

    # Add the results to the dataframe
    dtw_results_df_combined <- rbind(dtw_results_df_combined, 
                                     data.frame(DF_Name = df_names[i], # Use the dataframe name
                                                DTW_Component_1 = dtw_distance_1, 
                                                DTW_Component_2 = dtw_distance_2))
}

# Print and view the dataframe
print(dtw_results_df_combined)


```

Then we extract the relevant weights and add them together for the final dtw synchrony score:

```{r}
# Add columns for weighted DTW distances and their sum
dtw_results_df_combined$Weighted_DTW_Component_1 <- numeric(nrow(dtw_results_df_combined))
dtw_results_df_combined$Weighted_DTW_Component_2 <- numeric(nrow(dtw_results_df_combined))
dtw_results_df_combined$DTW_Sum <- numeric(nrow(dtw_results_df_combined))

# Loop through each row in the dtw_results_df_combined
for (i in 1:nrow(dtw_results_df_combined)) {
    # Extract the groupid from the name of the corresponding dataframe in component_list_A
    # This assumes the name format in component_list_A is something like "groupid_A"
    groupid <- sub("_A$", "", dtw_results_df_combined$DF_Name[i])

    # Find the corresponding row in PCA_component_weights
    matched_row <- PCA_component_weights[PCA_component_weights$groupid == groupid, ]
    
    # Extract the weight for the first dimension
    weight_1 <- matched_row$dimension_1_weight

    # Apply weights
    dtw_results_df_combined$Weighted_DTW_Component_1[i] <- dtw_results_df_combined$DTW_Component_1[i] * weight_1
    dtw_results_df_combined$Weighted_DTW_Component_2[i] <- dtw_results_df_combined$DTW_Component_2[i] * (1 - weight_1)

    # Calculate the sum
    dtw_results_df_combined$DTW_Sum[i] <- dtw_results_df_combined$Weighted_DTW_Component_1[i] + dtw_results_df_combined$Weighted_DTW_Component_2[i]
}

# Rename the dataframe for consistency
dtw_results_df <- dtw_results_df_combined

# Print the updated dataframe with the summed column
print(dtw_results_df)

# Saving it as dtw_results_df_PCA
write.csv(dtw_results_df, "results/dtw_results_df.csv", row.names = FALSE)


```

Calculating it manually to check whether the for loop calculated the DTW correctly.

## Visualising all pairs first principal component

All plots for the first and second component can be found in DTW_twowayPlots, DTW_threewayPlots and DTW_densityPlots in the results folder :)

### Twoway

```{r twoway for first component}
# Create empty vectors to store DTW distances, group numbers, and conditions
dtw_distances <- numeric(length(component_list_A))
group_numbers <- numeric(length(component_list_A))
conditions <- character(length(component_list_A))
# Create a folder to save plots
dir.create("results/DTW/DTW_twowayPlots", showWarnings = FALSE)

for (i in seq_along(component_list_A)) {
    group_number <- group_numbers[i]
    condition <- conditions[i]

    # Construct a meaningful title based on group name
    plot_title <- paste("First component for group", group_number, condition, "DTW")

    # Set the default font to Times New Roman
    par(family = "serif")

    # Open the PNG device
    png(file = paste0("results/DTW/DTW_twowayPlots/", gsub(" ", "_", plot_title), ".png"))

    # Plot
    plot(
        dtw(component_list_A[[i]][, "Dim.1"],
            component_list_B[[i]][, "Dim.1"],
            keep = TRUE),
        type = "twoway",
        lty = c(1, 1),
        lwd = c(6, 6),
        offset = -2,
        col = aesthetic_highlight_difference_palette,
        main = plot_title
    )

    # Close the PNG device after saving
    dev.off()
}

# Reset the default font after the loop
par(family = "sans")

```

### Threeway:

```{r threeway for first component}
# Create a folder to save plots
dir.create("results/DTW/DTW_threewayPlots", showWarnings = FALSE)

for (i in seq_along(component_list_A)) {
    group_number <- group_numbers[i]
    condition <- conditions[i]

    # Construct a meaningful title based on group name
    plot_title <- paste("First component for group", group_number, condition, "threeway")

    # Set the default font to Times New Roman
    par(family = "serif")

    # Open the PNG device
    png(file = paste0("results/DTW/DTW_threewayPlots/", gsub(" ", "_", plot_title), ".png"))

    # Plot
    plot(
        dtw(component_list_A[[i]][, "Dim.1"],
            component_list_B[[i]][, "Dim.1"],
            keep = TRUE),
       type = "threeway",
         off = 1,
         match.lty = 2,
        col = aesthetic_highlight_difference_palette,
        main = plot_title
    )

    # Close the PNG device after saving
    dev.off()
}

# Reset the default font after the loop
par(family = "sans")

```

### Density plots:

```{r Density plot for first component}
# Create a folder to save plots
dir.create("results/DTW/DTW_densityPlots", showWarnings = FALSE)

for (i in seq_along(component_list_A)) {
    group_number <- group_numbers[i]
    condition <- conditions[i]

    # Construct a meaningful title based on group name
    plot_title <- paste("First component for group", group_number, condition, "density")

    # Set the default font to Times New Roman
    par(family = "serif")

    # Open the PNG device
    png(file = paste0("results/DTW/DTW_densityPlots/", gsub(" ", "_", plot_title), ".png"))

    # Plot
    plot(
        dtw(component_list_A[[i]][, "Dim.1"],
            component_list_B[[i]][, "Dim.1"],
               k = TRUE),
         type = "density",
         off = 1,
         match.lty = 2,
        main = plot_title
    )

    # Close the PNG device after saving
    dev.off()
}

# Reset the default font after the loop
par(family = "sans")

```

## Visualising all pairs second principal component:

### Twoway:

```{r twoway for second component}
# Create a folder to save plots
dir.create("results/DTW/DTW_twowayPlots", showWarnings = FALSE)

for (i in seq_along(component_list_A)) {
    group_number <- group_numbers[i]
    condition <- conditions[i]

    # Construct a meaningful title based on group name
    plot_title <- paste("Second component for group", group_number, condition, "DTW")

    # Set the default font to Times New Roman
    par(family = "serif")

    # Open the PNG device
    png(file = paste0("results/DTW/DTW_twowayPlots/", gsub(" ", "_", plot_title), ".png"))

    # Plot
    plot(
        dtw(component_list_A[[i]][, "Dim.2"],
            component_list_B[[i]][, "Dim.2"],
            keep = TRUE),
        type = "twoway",
        lty = c(1, 1),
        lwd = c(6, 6),
        offset = -2,
        col = aesthetic_highlight_difference_palette,
        main = plot_title
    )

    # Close the PNG device after saving
    dev.off()
}

# Reset the default font after the loop
par(family = "sans")
```

### Threeway:

```{r threeway for second component}
# Create a folder to save plots
dir.create("results/DTW/DTW_threewayPlots", showWarnings = FALSE)

for (i in seq_along(component_list_A)) {
    group_number <- group_numbers[i]
    condition <- conditions[i]

    # Construct a meaningful title based on group name
    plot_title <- paste("Second component for group", group_number, condition, "threeway")

    # Set the default font to Times New Roman
    par(family = "serif")

    # Open the PNG device
    png(file = paste0("results/DTW/DTW_threewayPlots/", gsub(" ", "_", plot_title), ".png"))

    # Plot
    plot(
        dtw(component_list_A[[i]][, "Dim.2"],
            component_list_B[[i]][, "Dim.2"],
            keep = TRUE),
       type = "threeway",
         off = 1,
         match.lty = 2,
        col = aesthetic_highlight_difference_palette,
        main = plot_title
    )

    # Close the PNG device after saving
    dev.off()
}

# Reset the default font after the loop
par(family = "sans")

```

### Density:

```{r Density for second component}
# Create a folder to save plots
dir.create("results/DTW/DTW_densityPlots", showWarnings = FALSE)

for (i in seq_along(component_list_A)) {
    group_number <- group_numbers[i]
    condition <- conditions[i]

    # Construct a meaningful title based on group name
    plot_title <- paste("Second component for group", group_number, condition, "density")

    # Set the default font to Times New Roman
    par(family = "serif")

    # Open the PNG device
    png(file = paste0("results/DTW/DTW_densityPlots/", gsub(" ", "_", plot_title), ".png"))

    # Plot
    plot(
        dtw(component_list_A[[i]][, "Dim.2"],
            component_list_B[[i]][, "Dim.2"],
               k = TRUE),
         type = "density",
         off = 1,
         match.lty = 2,
        main = plot_title
    )

    # Close the PNG device after saving
    dev.off()
}

# Reset the default font after the loop
par(family = "sans")
```

## False-pairs

Having given each pair a score for their synchrony, we must now explore if this score is significant or not. We will do this by establishing false-pairs, looking at each subject A with all other subject B and scoring them. This will result in a spectrum of synchrony that is hopefully normally distributed. We will then see that the score for the correct pair is statistically significant. Because of limits in computational power, we will only do false pairs for one component.

We first look at the jointlead condition

```{r}
batch_size <- 5  # Set a reasonable batch size based on your memory
total_a <- sum(grepl("_jointlead_A$", names(component_list_A)))
total_b <- sum(grepl("_jointlead_B$", names(component_list_B)))
total_comparisons <- total_a * total_b

# Create a function to process a batch
process_batch <- function(start_idx, end_idx, component_list_A, component_list_B) {
    dtw_distances <- numeric(0)
    component_a_names <- character(0)
    component_b_names <- character(0)

    idx <- 1
    for (a_name in names(component_list_A)) {
        if (grepl("_jointlead_A$", a_name)) {
            component_a <- component_list_A[[a_name]]

            for (b_name in names(component_list_B)) {
                if (grepl("_jointlead_B$", b_name)) {
                    if (idx >= start_idx && idx <= end_idx) {
                        dtw_result <- dtw::dtw(component_a, component_list_B[[b_name]], keep = TRUE)
                        dtw_distances <- c(dtw_distances, dtw_result$distance)
                        component_a_names <- c(component_a_names, a_name)
                        component_b_names <- c(component_b_names, b_name)
                    }
                    idx <- idx + 1
                }
            }
        }
    }

    return(data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    ))
}

# Process in batches and combine results
results <- do.call(rbind, lapply(seq(1, total_comparisons, by = batch_size), function(x) {
    end_idx <- min(x + batch_size - 1, total_comparisons)
    process_batch(x, end_idx, component_list_A, component_list_B)
}))

# Create a data frame with results if any comparisons were made
if (length(dtw_distances) > 0) {
    dtw_results_df_jointlead <- data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    )
    print(dtw_results_df_jointlead)
} else {
    cat("No matching pairs found for '_leadfollow_A' and '_leadfollow_B'")
}


```

Then we look at the leadfollow condition:

```{r}
# Initialize vectors to store results
dtw_distances <- numeric(0)
component_a_names <- character(0)
component_b_names <- character(0)

# Loop through each element of component_list_A
for (a_name in names(component_list_A)) {
    if (grepl("_leadfollow_A$", a_name)) {
        component_a <- component_list_A[[a_name]]

        # Loop through each element of component_list_B for every qualifying component in component_list_A
        for (b_name in names(component_list_B)) {
            if (grepl("_leadfollow_B$", b_name)) {
                # Perform DTW comparison
                dtw_result <- dtw::dtw(component_a, component_list_B[[b_name]], keep = TRUE)
                
                # Store DTW distance and corresponding A and B component names
                dtw_distances <- c(dtw_distances, dtw_result$distance)
                component_a_names <- c(component_a_names, a_name)
                component_b_names <- c(component_b_names, b_name)
            }
        }
    }
}

# Create a data frame with results if any comparisons were made
if (length(dtw_distances) > 0) {
    dtw_results_df_leadfollow <- data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    )
    print(dtw_results_df_leadfollow)
} else {
    cat("No matching pairs found for '_leadfollow_A' and '_leadfollow_B'")
}

```

Now joining the two dataframes together

```{r}
# Combine the two dataframes
false_pairs_df <- rbind(dtw_results_df_leadfollow, dtw_results_df_jointlead)

rm(dtw_results_df_jointlead, dtw_results_df_leadfollow)
```

```{r}
# Loop over each unique value in the 'Component A' column
for (comp_a in unique(false_pairs_df$Component_A)) {
    # Filter the dataframe for the current component
    filtered_df <- subset(false_pairs_df, Component_A == comp_a)

    # Create a density plot
    p <- ggplot(filtered_df, aes(x = DTW_Distance)) +
        geom_density(fill = global_fill_colour, alpha = 0.5) +
      labs(
        title = paste("Density Plot of DTW Distances for Component A:"),
        subtitle = comp_a)

    # Display the plot
    print(p)
}

# Loop over each unique value in the 'Component B' column
for (comp_b in unique(false_pairs_df$Component_B)) {
    # Filter the dataframe for the current component
    filtered_df <- subset(false_pairs_df, Component_B == comp_b)

    # Create a density plot
    p <- ggplot(filtered_df, aes(x = DTW_Distance)) +
        geom_density(fill = global_fill_colour, alpha = 0.5) +
      labs(
        title = paste("Density Plot of DTW Distances for Component B:"),
        subtitle = comp_b)

    # Display the plot
    print(p)
}
```

## Real pairs

Taking the mean of DTW for all groups to get a baseline and checking which groups are below the mean and therefore high in similarity.

```{r}
DTW_mean <- mean(dtw_results_df$DTW_Distance) 

print(paste("mean:", DTW_mean))
for (i in seq_along(dtw_results_df$DTW_Distance)) {
  dtw_distance <- dtw_results_df$DTW_Distance[i]
  if (dtw_distance <= DTW_mean) {
    group <- dtw_results_df$Group[i]
    condition <- dtw_results_df$Condition[i]
    print(paste("DTW_Distance:", dtw_distance, "Group:", group, "Condition:", condition))
  }
}
```

```{r}
dtw_results_df %>%
  ggplot(aes(x = DTW_Sum)) +
  geom_density(fill = global_fill_colour, alpha = 0.5) +  # Add fill color and adjust transparency
  xlim(0,10000)+
  labs(
    title = "Density Plot of DTW Distances",
    x = "DTW Distance",
    y = "Density"
  ) 
```

# Exporting the synchrony scores

In order to compare this to each pairs empathy score, we export as a csv to be used in another RMD

```{r}
write_csv(dtw_results_df, "results/dtw_results_PCA.csv")
write_csv(false_pairs_df, "results/false_pairs_PCA.csv")
```
