---
title: "DTW ICA"
author: "Katharina Hellmund"
date: "2023-12-29"
output: pdf_document
---

```{r setup}
## Setup chunk
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)


pacman::p_load(
  "XML",
  "tidyverse",
  "fs",
  "assertthat",
  "stringi",
  "dtw",
  "RTransferEntropy",
  "signal",
  "conflicted",
  "Rcpp",
  "future",
  "dtw",
  "dplyr",
  "plotly",
  "htmlwidgets",
  "hrbrthemes",
  "zoo"
)
conflicts_prefer(dplyr::filter)
## Making sure we are in the right directory
wd <- getwd()
if (basename(wd) != "PerceptionActionExam") {
  setwd("./PerceptionActionExam")
}

                              ##NADIA PATH##
 # data_dir <- path_home() %>% 
 #   path("Documents", "GitHub", "PerceptionActionExam-Clean-up-attempt", "data", "tsvs") ## Nadia's path

                              #KATHARINA PATH##
data_dir <- path_home() %>%
path("Desktop","UNI", "3.semester", "Perception & Action", "PerceptionActionExam", "data", "tsvs") ## Katharina's path


#Aesthetic setup
theme_set(theme_ipsum(base_family = "Times New Roman"))
global_fill_colour <- "#8d5b5a"
aesthetic_palette <- c(
  "#d8aeb5","#c17f8c","#b59592","#9b6f69","#a94f62","#8d5b5a","#684141","#733545","#523438","#48222b","#2f1a1b")
aesthetic_highlight_difference_palette <- c("#d8aeb5","#2f1a1b")
```

# DTW on ICA

## Importing components

```{r}
# Directory where the components are stored
components_dir <- "results/ICA"

                  #IMPORTING ALL COMPONENTS FROM A#
# List all CSV files in the directory that include "_A_" in their names
component_files_A <- fs::dir_ls(path = components_dir, glob = "*_A_*.csv")

component_list_A <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in component_files_A) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  component_list_A[[file_name]] <- read.csv(csv_file)
}

                  #IMPORTING ALL COMPONENTS FROM B#
# List all CSV files in the directory that include "_B_" in their names
component_files_B <- fs::dir_ls(path = components_dir, glob = "*_B_*.csv")

component_list_B <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in component_files_B) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  component_list_B[[file_name]] <- read.csv(csv_file)
}
```

## Scoring each pair

Computing DTW for each group with a for loop, results can be found in "dtw_results_df"

```{r}
# Create empty vectors to store DTW distances, group numbers, and conditions
dtw_distances <- numeric(length(component_list_A))
group_numbers <- numeric(length(component_list_A))
conditions <- character(length(component_list_A))

# Loop through each pair of dataframes
for (i in seq_along(component_list_A)) {
  # Extract group number and condition from the element name for component_list_A
  matches <- regmatches(names(component_list_A)[i], regexec("group([0-9]+)_([^_]+)", names(component_list_A)[i]))
  group_number <- as.numeric(matches[[1]][2])
  condition <- matches[[1]][3]
  
  # Calculate DTW distance for component_list_A
  dtw_result <- dtw::dtw(component_list_A[[i]], component_list_B[[i]], keep = TRUE)
  
  # Extract DTW distance for component_list_A
  dtw_distance <- dtw_result$distance
  
  # Print the DTW distance for component_list_A
  print(paste("Group", group_number, condition, "DTW Distance:", dtw_distance))
  
  # Store DTW distance, group number, and condition in the vectors
  dtw_distances[i] <- dtw_distance
  group_numbers[i] <- group_number
  conditions[i] <- condition
}

# Create a data frame with results
dtw_results_df <- data.frame(Group = group_numbers, Condition = conditions, DTW_Distance = dtw_distances)

# Print overall summary or take further actions with dtw_distances vector
cat("Summary of DTW Distances:\n")
summary(dtw_distances)

# Display the data frame with results
print(dtw_results_df)

```

## Visualising

### Testing step patterns

```{r testing_step_patterns}
i<- 6
plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=rabinerJuangStepPattern(6,"c")),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)

plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=symmetricP1),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)
plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=symmetricP2),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)

plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=rabinerJuangStepPattern(7,"b")),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)
```

### Visualising all pairs:

```{r twoway}
for (i in seq_along(component_list_A)) {
    # TWO WAY plot
    plot(dtw(component_list_A[[i]],
             component_list_B[[i]],
             keep = TRUE,
             step = rabinerJuangStepPattern(6, "c")),
         type = "twoway",
         lty = c(1, 1),
         lwd = c(6, 6),
         offset = -2,
         col = aesthetic_highlight_difference_palette)

    # THREE WAY plot
    plot(dtw(component_list_A[[i]],
             component_list_B[[i]],
             k = TRUE),
         type = "threeway",
         off = 1,
         match.lty = 2,
         col = aesthetic_highlight_difference_palette)

    # DENSITY plot
    plot(dtw(component_list_A[[i]],
             component_list_B[[i]],
             k = TRUE),
         type = "density",
         off = 1,
         match.lty = 2)
}

```

This is a plot with a specificed step-pattern. I dont think we have to do this for the exam, as i saw Luuk not doing it and the differences between the step pattern seems quite hard to interpret on the internet. But to be fair, it looks pretty awesome, but there are a lot of other step-patterns out there and i am not sure, that this is the correct one to use for our data. Especially because this step pattern looks awesome with this group, but looks terrible with others.

Kat's answer:

I have found 3 different step patterns: Sakoe-chiba Band, Itakura Parallelogram and then Rabiner Juang (the one Luke is using), but it seems like you can define your own custom step pattern as well - that might be the most common? Anyways GPT says "The one you are currently using is typically used for speech recognition. It's good for sequences where one sequence may have more variance in timing or speed than the other."

## False-pairs

Having given each pair a score for their synchrony, we must now explore if this score is significant or not. We will do this by establishing false-pairs, looking at each subject A with all other subject B and scoring them. This will result in a spectrum of synchrony that is hopefully normally distrubuted. We will then see that the score for the correct pair is statistically significant.

We first look at the jointlead condition

```{r}
# Initialize vectors to store results
dtw_distances <- numeric(0)
component_a_names <- character(0)
component_b_names <- character(0)

# Loop through each element of component_list_A
for (a_name in names(component_list_A)) {
    if (grepl("_jointlead_A$", a_name)) {
        component_a <- component_list_A[[a_name]]

        # Loop through each element of component_list_B for every qualifying component in component_list_A
        for (b_name in names(component_list_B)) {
            if (grepl("_jointlead_B$", b_name)) {
                # Perform DTW comparison
                dtw_result <- dtw::dtw(component_a, component_list_B[[b_name]], keep = TRUE)
                
                # Store DTW distance and corresponding A and B component names
                dtw_distances <- c(dtw_distances, dtw_result$distance)
                component_a_names <- c(component_a_names, a_name)
                component_b_names <- c(component_b_names, b_name)
            }
        }
    }
}

# Create a data frame with results if any comparisons were made
if (length(dtw_distances) > 0) {
    dtw_results_df_jointlead <- data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    )
    print(dtw_results_df_jointlead)
} else {
    cat("No matching pairs found for '_jointlead_A' and '_jointlead_B'")
}

```

```{r}
# Initialize vectors to store results
dtw_distances <- numeric(0)
component_a_names <- character(0)
component_b_names <- character(0)

# Loop through each element of component_list_A
for (a_name in names(component_list_A)) {
    if (grepl("_leadfollow_A$", a_name)) {
        component_a <- component_list_A[[a_name]]

        # Loop through each element of component_list_B for every qualifying component in component_list_A
        for (b_name in names(component_list_B)) {
            if (grepl("_leadfollow_B$", b_name)) {
                # Perform DTW comparison
                dtw_result <- dtw::dtw(component_a, component_list_B[[b_name]], keep = TRUE)
                
                # Store DTW distance and corresponding A and B component names
                dtw_distances <- c(dtw_distances, dtw_result$distance)
                component_a_names <- c(component_a_names, a_name)
                component_b_names <- c(component_b_names, b_name)
            }
        }
    }
}

# Create a data frame with results if any comparisons were made
if (length(dtw_distances) > 0) {
    dtw_results_df_leadfollow <- data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    )
    print(dtw_results_df_leadfollow)
} else {
    cat("No matching pairs found for '_leadfollow_A' and '_leadfollow_B'")
}

```

Now joining the two dataframes together

```{r}
# Combine the two dataframes
false_pairs_df <- rbind(dtw_results_df_leadfollow, dtw_results_df_jointlead)

rm(dtw_results_df_jointlead, dtw_results_df_leadfollow)
```

```{r}
# Loop over each unique value in the 'Component A' column
for (comp_a in unique(false_pairs_df$Component_A)) {
    # Filter the dataframe for the current component
    filtered_df <- subset(false_pairs_df, Component_A == comp_a)

    # Create a density plot
    p <- ggplot(filtered_df, aes(x = DTW_Distance)) +
        geom_density(fill = global_fill_colour, alpha = 0.5) +
      labs(
        title = paste("Density Plot of DTW Distances for Component A:"),
        subtitle = comp_a)

    # Display the plot
    print(p)
}

# Loop over each unique value in the 'Component B' column
for (comp_b in unique(false_pairs_df$Component_B)) {
    # Filter the dataframe for the current component
    filtered_df <- subset(false_pairs_df, Component_B == comp_b)

    # Create a density plot
    p <- ggplot(filtered_df, aes(x = DTW_Distance)) +
        geom_density(fill = global_fill_colour, alpha = 0.5) +
      labs(
        title = paste("Density Plot of DTW Distances for Component B:"),
        subtitle = comp_b)

    # Display the plot
    print(p)
}
```

## Real pairs

Taking the mean of DTW for all groups to get a baseline and checking which groups are below the mean and therefore high in similarity.

```{r}
DTW_mean <- mean(dtw_results_df$DTW_Distance) 

print(paste("mean:", DTW_mean))
for (i in seq_along(dtw_results_df$DTW_Distance)) {
  dtw_distance <- dtw_results_df$DTW_Distance[i]
  if (dtw_distance <= DTW_mean) {
    group <- dtw_results_df$Group[i]
    condition <- dtw_results_df$Condition[i]
    print(paste("DTW_Distance:", dtw_distance, "Group:", group, "Condition:", condition))
  }
}
```

```{r}
dtw_results_df %>%
  ggplot(aes(x = DTW_Distance)) +
  geom_density(fill = global_fill_colour, alpha = 0.5) +  # Add fill color and adjust transparency
  labs(
    title = "Density Plot of DTW Distances",
    x = "DTW Distance",
    y = "Density"
  ) 
```

# Exporting the synchrony scores

In order to compare this to each pairs empathy score, we export as a csv to be used in another RMD

```{r}
write_csv(dtw_results_df, "results/dtw_results_ICA.csv")
```
