---
title: "DTW_PCA"
author: "Katharina Hellmund"
date: "2023-12-29"
output: pdf_document
---

```{r setup}
## Setup chunk
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)


pacman::p_load(
  "XML",
  "tidyverse",
  "fs",
  "assertthat",
  "stringi",
  "dtw",
  "RTransferEntropy",
  "signal",
  "conflicted",
  "Rcpp",
  "future",
  "dtw",
  "dplyr",
  "plotly",
  "htmlwidgets",
  "hrbrthemes",
  "zoo"
)
conflicts_prefer(dplyr::filter)
## Making sure we are in the right directory
wd <- getwd()
if (basename(wd) != "PerceptionActionExam") {
  setwd("./PerceptionActionExam")
}

                              ##NADIA PATH##
  data_dir <- path_home() %>% 
   path("Documents", "GitHub", "PerceptionActionExam-Clean-up-attempt", "data", "tsvs") ## Nadia's path

                              #KATHARINA PATH##
 #data_dir <- path_home() %>%
 # path("Desktop","UNI", "3.semester", "Perception & Action", "PerceptionActionExam", "data", "tsvs") ## Katharina's path


#Aesthetic setup
theme_set(theme_ipsum(base_family = "Times New Roman"))
global_fill_colour <- "#8d5b5a"
aesthetic_palette <- c(
  "#d8aeb5","#c17f8c","#b59592","#9b6f69","#a94f62","#8d5b5a","#684141","#733545","#523438","#48222b","#2f1a1b")
aesthetic_highlight_difference_palette <- c("#d8aeb5","#2f1a1b")
```

# DTW on PCA

## Importing components

```{r}
# Directory where the components are stored
components_dir <- "results/PCA"

                  #IMPORTING ALL COMPONENTS FROM A#
# List all CSV files in the directory that include "_A" in their names
component_files_A <- fs::dir_ls(path = components_dir, glob = "*_A.csv")

component_list_A <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in component_files_A) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  component_list_A[[file_name]] <- read.csv(csv_file)
}

                  #IMPORTING ALL COMPONENTS FROM B#
# List all CSV files in the directory that include "_B" in their names
component_files_B <- fs::dir_ls(path = components_dir, glob = "*_B.csv")

component_list_B <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in component_files_B) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  component_list_B[[file_name]] <- read.csv(csv_file)
}
```

## Scoring each pair (One component)

Computing DTW for each group with a for loop, results can be found in "dtw_results_df"

```{r}
# Create empty vectors to store DTW distances, group numbers, and conditions
dtw_distances <- numeric(length(component_list_A))
group_numbers <- numeric(length(component_list_A))
conditions <- character(length(component_list_A))

# Loop through each pair of dataframes
for (i in seq_along(component_list_A)) {
  # Extract group number and condition from the element name for component_list_A
  matches <- regmatches(names(component_list_A)[i], regexec("group([0-9]+)_([^_]+)", names(component_list_A)[i]))
  group_number <- as.numeric(matches[[1]][2])
  condition <- matches[[1]][3]
  
  # Calculate DTW distance for component_list_A
  dtw_result <- dtw::dtw(component_list_A[[i]], component_list_B[[i]], keep = TRUE)
  
  # Extract DTW distance for component_list_A
  dtw_distance <- dtw_result$distance
  
  # Print the DTW distance for component_list_A
  print(paste("Group", group_number, condition, "DTW Distance:", dtw_distance))
  
  # Store DTW distance, group number, and condition in the vectors
  dtw_distances[i] <- dtw_distance
  group_numbers[i] <- group_number
  conditions[i] <- condition
}

# Create a data frame with results
dtw_results_df <- data.frame(Group = group_numbers, Condition = conditions, DTW_Distance = dtw_distances)

# Print overall summary or take further actions with dtw_distances vector
cat("Summary of DTW Distances:\n")
summary(dtw_distances)

# Display the data frame with results
print(dtw_results_df)
```
## NEW: Scoring each pair (Two components)
```{r}
# Create empty vectors to store DTW distances, group numbers, and conditions
dtw_distances <- numeric(length(component_list_A))
group_numbers <- numeric(length(component_list_A))
conditions <- character(length(component_list_A))

# Loop through each pair of dataframes
for (i in seq_along(component_list_A)) {
  # Extract group number and condition from the element name for component_list_A
  matches <- regmatches(names(component_list_A)[i], regexec("group([0-9]+)_([^_]+)", names(component_list_A)[i]))
  group_number <- as.numeric(matches[[1]][2])
  condition <- matches[[1]][3]
  
  # Extract the first principal component for component_list_A and component_list_B
  component_A_1 <- component_list_A[[i]][, "Dim.1"]
  component_B_1 <- component_list_B[[i]][, "Dim.1"]
  
  # Extract the second principal component for component_list_A and component_list_B
  component_A_2 <- component_list_A[[i]][, "Dim.2"]
  component_B_2 <- component_list_B[[i]][, "Dim.2"]
  
  # Calculate DTW distance for the first principal component
  dtw_result1 <- dtw::dtw(component_A_1, component_B_1, keep = TRUE)
  
  # Extract DTW distance for the first principal component
  dtw_distance1 <- dtw_result1$distance
  
  # Calculate DTW distance for the second principal component
  dtw_result2 <- dtw::dtw(component_A_2, component_B_2, keep = TRUE)
  
  # Extract DTW distance for the second principal component
  dtw_distance2 <- dtw_result2$distance
  
  # Sum the DTW distances for the first and second components
  dtw_together <- dtw_distance1 + dtw_distance2
  
  # Divide the summed DTW distance by 2 (take the mean)
  dtw_distance_mean <- dtw_together / 2
  
  # Print the mean DTW distance for both components
  print(paste("Group", group_number, condition, "DTW Distance:", dtw_distance_mean))
  
  # Store mean DTW distance, group number, and condition in the vectors
  dtw_distances[i] <- dtw_distance_mean
  group_numbers[i] <- group_number
  conditions[i] <- condition
}

# Create a data frame with results
dtw_results_df_PCA <- data.frame(Group = group_numbers, Condition = conditions, Mean_DTW_Distance = dtw_distances)

# Print overall summary or take further actions with dtw_distances vector
cat("Summary of Mean DTW Distances:\n")
summary(dtw_distances)

# Display the data frame with results
print(dtw_results_df_PCA)

# Saving it as dtw_results_df_PCA
write_csv(dtw_results_df_PCA, "results/dtw_results_df_PCA.csv")
```


## NEW: Calculating it manually to check whether the for loop calculated the DTW correctly (It did wuhuu!)
```{r}
## extracting the principal components
A <- component_list_A[["group0_jointlead_A"]]
B <- component_list_B[["group0_jointlead_B"]]

## first principal component
dim1 <- dtw(A$Dim.1, B$Dim.1, keep = TRUE)

## calculating distance for first principal component
dim1_distance <- dim1$distance

## second principal component
dim2 <- dtw(A$Dim.2, B$Dim.2, keep = TRUE)

## calculating distance for second principal component
dim2_distance <- dim2$distance

## final DTW result
final <- (dim1_distance + dim2_distance )/2

print(final)
```

## Visualising

### Testing step patterns

```{r testing_step_patterns}
i<- 6
plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=rabinerJuangStepPattern(6,"c")),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)

plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=symmetricP1),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)
plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=symmetricP2),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)

plot(
    dtw(component_list_A[[i]],component_list_B[[i]],
        keep=TRUE,
        step=rabinerJuangStepPattern(7,"b")),
        col = aesthetic_highlight_difference_palette,
         lty = c(1, 1),
         lwd = c(6, 6),
        type="twoway",offset=-2)
```

### Visualising all pairs: (this is old)

```{r twoway}
for (i in seq_along(component_list_A)) {
    # TWO WAY plot
    plot(dtw(component_list_A[[i]],
             component_list_B[[i]],
             keep = TRUE),
         type = "twoway",
         lty = c(1, 1),
         lwd = c(6, 6),
         offset = -2,
         col = aesthetic_highlight_difference_palette)

    # THREE WAY plot
    plot(dtw(component_list_A[[i]],
             component_list_B[[i]],
             k = TRUE),
         type = "threeway",
         off = 1,
         match.lty = 2,
         col = aesthetic_highlight_difference_palette)

    # DENSITY plot
    plot(dtw(component_list_A[[i]],
             component_list_B[[i]],
             k = TRUE),
         type = "density",
         off = 1,
         match.lty = 2)
}

```

### NEW: Visualising all pairs first principal component:

```{r twoway}
for (i in seq_along(component_list_A)) {
    # TWO WAY plot
    plot(dtw(component_list_A[[i]][, "Dim.1"],
             component_list_B[[i]][, "Dim.1"],
             keep = TRUE),
         type = "twoway",
         lty = c(1, 1),
         lwd = c(6, 6),
         offset = -2,
         col = aesthetic_highlight_difference_palette)

    # THREE WAY plot
    plot(dtw(component_list_A[[i]][, "Dim.1"],
             component_list_B[[i]][, "Dim.1"],
             k = TRUE),
         type = "threeway",
         off = 1,
         match.lty = 2,
         col = aesthetic_highlight_difference_palette)

    # DENSITY plot
    plot(dtw(component_list_A[[i]][, "Dim.1"],
             component_list_B[[i]][, "Dim.1"],
             k = TRUE),
         type = "density",
         off = 1,
         match.lty = 2)
}

```

### NEW: Visualising all pairs second principal component:

```{r twoway}
for (i in seq_along(component_list_A)) {
    # TWO WAY plot
   plot(dtw(component_list_A[[i]][, "Dim.2"],
             component_list_B[[i]][, "Dim.2"],
             keep = TRUE),
         type = "twoway",
         lty = c(1, 1),
         lwd = c(6, 6),
         offset = -2,
         col = aesthetic_highlight_difference_palette)

    # THREE WAY plot
 plot(dtw(component_list_A[[i]][, "Dim.2"],
             component_list_B[[i]][, "Dim.2"],
             k = TRUE),
         type = "threeway",
         off = 1,
         match.lty = 2,
         col = aesthetic_highlight_difference_palette)

    # DENSITY plot
     plot(dtw(component_list_A[[i]][, "Dim.2"],
             component_list_B[[i]][, "Dim.2"],
             k = TRUE),
         type = "density",
         off = 1,
         match.lty = 2)
}

```



## False-pairs

Having given each pair a score for their synchrony, we must now explore if this score is significant or not. We will do this by establishing false-pairs, looking at each subject A with all other subject B and scoring them. This will result in a spectrum of synchrony that is hopefully normally distrubuted. We will then see that the score for the correct pair is statistically significant.

We first look at the jointlead condition

```{r}
# Initialize vectors to store results
dtw_distances <- numeric(0)
component_a_names <- character(0)
component_b_names <- character(0)

# Loop through each element of component_list_A
for (a_name in names(component_list_A)) {
    if (grepl("_jointlead_A$", a_name)) {
        component_a <- component_list_A[[a_name]]

        # Loop through each element of component_list_B for every qualifying component in component_list_A
        for (b_name in names(component_list_B)) {
            if (grepl("_jointlead_B$", b_name)) {
                # Perform DTW comparison
                dtw_result <- dtw::dtw(component_a, component_list_B[[b_name]], keep = TRUE)
                
                # Store DTW distance and corresponding A and B component names
                dtw_distances <- c(dtw_distances, dtw_result$distance)
                component_a_names <- c(component_a_names, a_name)
                component_b_names <- c(component_b_names, b_name)
            }
        }
    }
}

# Create a data frame with results if any comparisons were made
if (length(dtw_distances) > 0) {
    dtw_results_df_jointlead <- data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    )
    print(dtw_results_df_jointlead)
} else {
    cat("No matching pairs found for '_jointlead_A' and '_jointlead_B'")
}

```

```{r}
# Initialize vectors to store results
dtw_distances <- numeric(0)
component_a_names <- character(0)
component_b_names <- character(0)

# Loop through each element of component_list_A
for (a_name in names(component_list_A)) {
    if (grepl("_leadfollow_A$", a_name)) {
        component_a <- component_list_A[[a_name]]

        # Loop through each element of component_list_B for every qualifying component in component_list_A
        for (b_name in names(component_list_B)) {
            if (grepl("_leadfollow_B$", b_name)) {
                # Perform DTW comparison
                dtw_result <- dtw::dtw(component_a, component_list_B[[b_name]], keep = TRUE)
                
                # Store DTW distance and corresponding A and B component names
                dtw_distances <- c(dtw_distances, dtw_result$distance)
                component_a_names <- c(component_a_names, a_name)
                component_b_names <- c(component_b_names, b_name)
            }
        }
    }
}

# Create a data frame with results if any comparisons were made
if (length(dtw_distances) > 0) {
    dtw_results_df_leadfollow <- data.frame(
        Component_A = component_a_names, 
        Component_B = component_b_names, 
        DTW_Distance = dtw_distances
    )
    print(dtw_results_df_leadfollow)
} else {
    cat("No matching pairs found for '_leadfollow_A' and '_leadfollow_B'")
}

```

Now joining the two dataframes together

```{r}
# Combine the two dataframes
false_pairs_df <- rbind(dtw_results_df_leadfollow, dtw_results_df_jointlead)

rm(dtw_results_df_jointlead, dtw_results_df_leadfollow)
```

```{r}
# Loop over each unique value in the 'Component A' column
for (comp_a in unique(false_pairs_df$Component_A)) {
    # Filter the dataframe for the current component
    filtered_df <- subset(false_pairs_df, Component_A == comp_a)

    # Create a density plot
    p <- ggplot(filtered_df, aes(x = DTW_Distance)) +
        geom_density(fill = global_fill_colour, alpha = 0.5) +
      labs(
        title = paste("Density Plot of DTW Distances for Component A:"),
        subtitle = comp_a)

    # Display the plot
    print(p)
}

# Loop over each unique value in the 'Component B' column
for (comp_b in unique(false_pairs_df$Component_B)) {
    # Filter the dataframe for the current component
    filtered_df <- subset(false_pairs_df, Component_B == comp_b)

    # Create a density plot
    p <- ggplot(filtered_df, aes(x = DTW_Distance)) +
        geom_density(fill = global_fill_colour, alpha = 0.5) +
      labs(
        title = paste("Density Plot of DTW Distances for Component B:"),
        subtitle = comp_b)

    # Display the plot
    print(p)
}
```

## Real pairs

Taking the mean of DTW for all groups to get a baseline and checking which groups are below the mean and therefore high in similarity.

```{r}
DTW_mean <- mean(dtw_results_df$DTW_Distance) 

print(paste("mean:", DTW_mean))
for (i in seq_along(dtw_results_df$DTW_Distance)) {
  dtw_distance <- dtw_results_df$DTW_Distance[i]
  if (dtw_distance <= DTW_mean) {
    group <- dtw_results_df$Group[i]
    condition <- dtw_results_df$Condition[i]
    print(paste("DTW_Distance:", dtw_distance, "Group:", group, "Condition:", condition))
  }
}
```

```{r}
dtw_results_df %>%
  ggplot(aes(x = DTW_Distance)) +
  geom_density(fill = global_fill_colour, alpha = 0.5) +  # Add fill color and adjust transparency
  labs(
    title = "Density Plot of DTW Distances",
    x = "DTW Distance",
    y = "Density"
  ) 
```

# Exporting the synchrony scores

In order to compare this to each pairs empathy score, we export as a csv to be used in another RMD

```{r}
write_csv(dtw_results_df, "results/dtw_results_PCA.csv")
write_csv(false_pairs_df, "results/false_pairs_PCA.csv")
```
