---
title: "From raw motion-data to component:ICA and PCA"
author: "Katharina Hellmund"
date: "2023-12-29"
output: pdf_document
---

```{r setup}
## Setup chunk
knitr::opts_chunk$set(echo = TRUE, include = TRUE, message = FALSE, warning = FALSE)


pacman::p_load(
  "XML",
  "tidyverse",
  "fs",
  "assertthat",
  "stringi",
  "dtw",
  "RTransferEntropy",
  "signal",
  "conflicted",
  "Rcpp",
  "future",
  "dtw",
  "dplyr",
  "plotly",
  "htmlwidgets",
  "hrbrthemes",
  "zoo"
)
conflicts_prefer(dplyr::filter)
## Making sure we are in the right directory
wd <- getwd()
if (basename(wd) != "PerceptionActionExam") {
  setwd("./PerceptionActionExam")
}

                              ##NADIA PATH##
 # data_dir <- path_home() %>% 
 #   path("Documents", "GitHub", "PerceptionActionExam-Clean-up-attempt", "data", "tsvs") ## Nadia's path

                              #KATHARINA PATH##
data_dir <- path_home() %>%
path("Desktop","UNI", "3.semester", "Perception & Action", "PerceptionActionExam", "data", "tsvs") ## Katharina's path


#Aesthetic setup
theme_set(theme_ipsum(base_family = "Times New Roman"))
global_fill_colour <- "#8d5b5a"
aesthetic_palette <- c(
  "#d8aeb5","#c17f8c","#b59592","#9b6f69","#a94f62","#8d5b5a","#684141","#733545","#523438","#48222b","#2f1a1b")
aesthetic_highlight_difference_palette <- c("#d8aeb5","#2f1a1b")
```

# Importing Data

We import the preprocessed data

```{r}
# Set the directory where your CSV files are located
data_folder <- "data/mocap_data_prepped"

# List all CSV files in the specified folder
csv_files <- list.files(data_folder, pattern = "*.csv", full.names = TRUE)

# Initialize an empty list to store the dataframes
list_of_dataframes <- list()

# Loop through each CSV file, read it into a dataframe, and store it in the list with its original file name
for (csv_file in csv_files) {
  # Extract the file name without the extension
  file_name <- tools::file_path_sans_ext(basename(csv_file))
  
  # Read the CSV file into a dataframe and assign it to the list with the file name as the list name
  list_of_dataframes[[file_name]] <- read.csv(csv_file)
}

```

# ICA: Independent Component Analysis

## ICA-preprocessing

In order for ICA to work, our data must be exclusively numeric, meaning we will need to prep dataframes to include only our coordinates. The goal is to get a measure of the subjects movement per time, therefore we split our data into two: A and B. We also omit any NAs. This is subpar, but we have decided it is better to omit NAs entirely rather than apply an average throughout the missing rows - as this might skew it in favour of looking more synchronised.

```{r}
list_of_dataframes_A <- list()
list_of_dataframes_B <- list()

# Get the names of the original dataframes
original_names <- names(list_of_dataframes)

for (i in seq_along(list_of_dataframes)) {
    df <- list_of_dataframes[[i]]
    df_name <- original_names[i]

    # Split the dataframe based on 'subject' and omit rows with NAs
    df_A <- na.omit(df[df$subject == 'A', ])
    df_B <- na.omit(df[df$subject == 'B', ])

    # Add the split and cleaned dataframes to their respective lists with modified names
    if(nrow(df_A) > 0) {
        list_of_dataframes_A[[paste(df_name, "A", sep = "_")]] <- df_A
    }
    if(nrow(df_B) > 0) {
        list_of_dataframes_B[[paste(df_name, "B", sep = "_")]] <- df_B
    }
}
rm(df, df_A, df_B, i, df_name)
```

Now that we have split our data, we need to convert it into a matrix with xyz coordinates where each row is a time-frame. In order to do this, we need to go back to a wider format, where each row is a single time point. This means each marker x y z will take up columns. We exclude elapsed_time, because it will be represented row-by-row. We also need to ungroup the dataframe, otherwise we can't remove columns. Furthermore, once changed to matrix, we have no choice but to impute NA or discard all rows with NAs, which is not viable.

```{r}
# Define a function to replace NAs with column means
replace_nas_with_column_means <- function(df) {
  df %>%
    mutate(across(everything(), ~ifelse(is.na(.), mean(., na.rm = TRUE), .)))
}

# Iterate over list_of_dataframes_A and replace NAs with column means
for (i in seq_along(list_of_dataframes_A)) {
  list_of_dataframes_A[[i]] <- list_of_dataframes_A[[i]] %>%
    ungroup() %>%
    select(-condition, -index, -subject, -group) %>%
    gather(key, value, -elapsed_time, -marker) %>%
    unite(col = "new_col", marker, key, remove = TRUE) %>%
    spread(key = new_col, value = value) %>%
    arrange(elapsed_time) %>%
    replace_nas_with_column_means()
}

# Iterate over list_of_dataframes_B and replace NAs with column means
for (i in seq_along(list_of_dataframes_B)) {
  list_of_dataframes_B[[i]] <- list_of_dataframes_B[[i]] %>%
    ungroup() %>%
    select(-condition, -index, -subject, -group) %>%
    gather(key, value, -elapsed_time, -marker) %>%
    unite(col = "new_col", marker, key, remove = TRUE) %>%
    spread(key = new_col, value = value) %>%
    arrange(elapsed_time) %>%
    replace_nas_with_column_means()
}
rm(i)
```

Now we are ready to convert to matrix, we also standardize as a step in this function with scale(). ICA often benefit from having its features on the same scale, and therefore standardising before running ICA is good practise.

```{r}
# Function to convert a dataframe to a matrix with standardized values for ICA
convert_df_to_matrix_for_ICA <- function(df) {
    # Ensure the dataframe is sorted by elapsed_time
    df_sorted <- df %>% arrange(elapsed_time)

    # Select only numeric columns
    df_numeric <- df_sorted %>% select(where(is.numeric))

    # Standardize the data
    df_standardized <- as.data.frame(scale(df_numeric))

    # Return as matrix
    return(as.matrix(df_standardized))
}

# Convert each dataframe in list_of_dataframes_A to a matrix for ICA
matrices_A <- lapply(list_of_dataframes_A, convert_df_to_matrix_for_ICA)

# Convert each dataframe in list_of_dataframes_B to a matrix for ICA
matrices_B <- lapply(list_of_dataframes_B, convert_df_to_matrix_for_ICA)
```

```{r}
# Check for NAs in matrices_A
any(is.na(unlist(matrices_A)))

# Check for NAs in matrices_B
any(is.na(unlist(matrices_B)))

```

## Deciding on amount of components

We want to make sure we select the optimal amount of components to describe each subjects motion.

```{r}

```

## Running ICA

Though we may lose some data-complexity, we decide to extract only 1 component. Furthermore, ICA sometimes flips data as it does not inherently know which direction the data faces. This is a big issue when using ICA before dynamtic time warping, as we need to make sure both components A and B have the same direction.

Since we know the components represent synchrony we compute the area between the components in all configurations, and choose the configuration where the area is the smallest. In other words, the components will be flipped in a way to consistently be close to eachother and therefore this should result in them being faced "the same way". This method only works because we know the nature of the components is to attempt to follow eachother for A and B.

We will discuss using ICA with motion capture data as a method in the discussion.

First we apply ICA on the matrix:

```{r}
# Function to apply ICA on a matrix
apply_ica <- function(matrix_data, n_comp) {
  # Assuming the first column is 'elapsed_time' and should not be included in ICA
  matrix_for_ica <- matrix_data[, -1]  # Exclude 'elapsed_time'
  
  # Run ICA (nAdiA:I CHANGED ROW.NORM)
  if(ncol(matrix_for_ica) >= n_comp) {
    ica_result <- fastICA(matrix_for_ica, n.comp = n_comp,alg.typ = "parallel", fun = "logcosh", alpha = 1,
             method = "R", row.norm = TRUE, maxit = 200,
             tol = 0.0001, verbose = TRUE)
    return(ica_result)
  } else {
    warning(paste("Not enough columns in matrix to extract", n_comp, "components"))
    return(NULL)
  }
}

# Apply ICA on matrices in list_of_dataframes_A
ica_results_A <- lapply(matrices_A, apply_ica, n_comp = 1)  # Adjust n.comp as needed


# Apply ICA on matrices in list_of_dataframes_B
ica_results_B <- lapply(matrices_B, apply_ica, n_comp = 1)  # Adjust n.comp as needed

component_list_A <- lapply(ica_results_A, function(ica) ica$S)
component_list_B <- lapply(ica_results_B, function(ica) ica$S)

rm(ica_results_A, ica_results_B)
```

Next we calculate the area between the components and pick the combination which results in the smallest area: in our case that must be the case where they are aligned.

```{r}
# Function to calculate the area between two components, aligning their lengths
calculate_area_between_components <- function(comp1, comp2) {
  # Align lengths
  min_length <- min(nrow(comp1), nrow(comp2))
  comp1_aligned <- comp1[1:min_length, ]
  comp2_aligned <- comp2[1:min_length, ]

  # Calculate area
  sum(abs(comp1_aligned - comp2_aligned))
}

# Function to align a component based on the smallest area between lines
align_component_based_on_area <- function(component_A, component_B) {
  # Align component lengths for area calculation
  original_area <- calculate_area_between_components(component_A, component_B)
  flipped_area <- calculate_area_between_components(component_A, -component_B)

  if (flipped_area < original_area) {
    return(-component_B)
  } else {
    return(component_B)
  }
}

# Align each pair of components from A and B
for (i in seq_along(component_list_A)) {
  if (!is.null(component_list_A[[i]]) && !is.null(component_list_B[[i]])) {
    component_list_B[[i]] <- align_component_based_on_area(component_list_A[[i]], component_list_B[[i]])
  }
}


# Clean up
rm(list_of_dataframes_A, list_of_dataframes_B)

```

Now we visualise the components per condition in a common plot, to make it clear which representations of motion each subject has, effectively making it easier to visualise their pathways as a last step before DTW.

```{r}
# Loop through each index of the component lists
for (i in seq_along(component_list_A)) {
  # Extract components for Subject A and Subject B
  component_A <- component_list_A[[i]]
  component_B <- component_list_B[[i]]

  # Ensure both components have the same number of rows
  min_rows <- min(nrow(component_A), nrow(component_B))

  # Combine both components into a single dataframe
  combined_df <- data.frame(
    time = rep(1:min_rows, 2),
    value = c(component_A[1:min_rows, 1], component_B[1:min_rows, 1]),
    group = rep(c("Subject A", "Subject B"), each = min_rows)
  )

  # Create a time series plot for both subjects
  component_name_A <- names(component_list_A)[i]
  component_name_B <- names(component_list_B)[i]
  group_number <- paste(component_name_A, "vs", component_name_B)

  
  p <- ggplot(combined_df, aes(x = time, y = value, color = group)) +
    geom_line(size = 1) +
    labs(
      title = "Independent components", 
      subtitle = group_number,
         x = "Time", y = "Component Value", color = "Subject") +
    scale_color_manual(values = aesthetic_highlight_difference_palette)

  # Print the plot
  print(p)
}


```

OBS: CHECK GROUP 12 LEADFOLLOW, GROUP 13 JOINTLEAD & LEADFOLLOW, GROUP 6 JOINTLEAD, -\> ARE THE COMPONENTS NOT REFLECTIVE OF WHATS GOING ON OR ARE THEY JUST REALLY BAD AT JOINT MOTION.

## Saving the components

We will now save the components to be tested with dynamic time warping in a seperate markdown. We will save both the lists of components in respective folders

```{r}
library(fs)

# Create the "results/ICA" directory if it doesn't already exist
dir_create(path("results", "ICA"), recursive = TRUE)

# Function to save each matrix in the component list as a CSV
save_component_list_as_csv <- function(component_list, dir_path) {
  # Ensure component_list has named indices
  if (is.null(names(component_list))) {
    stop("The component list must have named indices.")
  }

  for (comp_name in names(component_list)) {
    # Use the component name for the file name
    file_path <- file.path(dir_path, paste0(comp_name, ".csv"))
    write.csv(component_list[[comp_name]], file_path, row.names = FALSE)
  }

  # Print a message to indicate completion
  cat("All components have been saved as CSV in the '", dir_path, "' directory.\n")
}

# Example usage: Save component_list_A and component_list_B to "results/ICA" as CSV files
# Ensure that component_list_A and component_list_B have named indices
save_component_list_as_csv(component_list_A, "results/ICA")
save_component_list_as_csv(component_list_B, "results/ICA")

```
